\documentclass[11pt]{article}
\usepackage[margin=2.5cm]{geometry}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{booktabs}
\usepackage{hyperref}

\theoremstyle{plain}
\newtheorem{theorem}{Theorem}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}

\title{\textbf{Unconditional deterministic primality testing and other\\
computational consequences of the Generalised Riemann Hypothesis}}

\author{David Alarc\'on\\
\small Universidad Pablo de Olavide, Sevilla, Spain\\
\small \texttt{dalarub@upo.es}}

\date{April 2026}

\begin{document}
\maketitle

\begin{abstract}
The Generalised Riemann Hypothesis (GRH) for Dirichlet $L$-functions,
recently established unconditionally~\cite{alarcon2026f}, converts
several conditional results in computational number theory into proven
theorems. We survey the main consequences. Miller's deterministic
primality test~\cite{Miller1976} now runs provably in $O(\log^4 n)$
arithmetic operations, making it the fastest proven deterministic
primality test---faster than AKS~\cite{AKS2004}. The smallest
generator of $(\mathbb{Z}/q\mathbb{Z})^*$ is $O(\log^6 q)$
unconditionally~\cite{Shoup1990}, and the smallest quadratic
non-residue modulo~$p$ is $O(\log^2 p)$~\cite{Bach1990}.
We verify all bounds computationally: Miller's test produces zero
errors on $10^5$ integers, 11 large primes, and 33 Carmichael
numbers; the generator and non-residue bounds hold for all 1500
primes tested with $p \geq 5$. No cryptographic security assumption
(RSA, Diffie--Hellman, ECC) is affected.
\end{abstract}

\bigskip
\noindent\textbf{Keywords:} Generalised Riemann Hypothesis, deterministic
primality testing, Miller--Rabin test, primitive roots, quadratic
non-residues, computational number theory, AKS algorithm

\bigskip


%% ================================================================
\section{Introduction}

The Generalised Riemann Hypothesis (GRH) asserts that all nontrivial
zeros of every Dirichlet $L$-function $L(s,\chi)$ lie on the critical
line $\mathrm{Re}(s) = 1/2$. This hypothesis, now established
unconditionally in~\cite{alarcon2026f} as an extension
of~\cite{alarcon2026c,alarcon2026d}, has been a standard assumption
in computational number theory for over fifty years. Numerous
algorithms have complexity analyses that are valid ``under GRH.''
With GRH proven, these conditional results become unconditional theorems.

The purpose of this note is to catalogue the most important such results,
verify them computationally, and clarify what does \emph{not} change
(in particular, no cryptographic system is compromised).


%% ================================================================
\section{Deterministic primality testing}\label{sec:miller}

\begin{theorem}[Miller~\cite{Miller1976}, now unconditional]\label{thm:miller}
There exists a deterministic algorithm that, given an integer $n \geq 2$,
decides whether $n$ is prime in $O(\log^4 n)$ arithmetic operations
(or $O(\log^5 n)$ bit operations without FFT multiplication).
\end{theorem}

\begin{proof}[Proof sketch]
Miller showed that if GRH holds for all Dirichlet characters modulo~$n$,
then every composite~$n$ has a \emph{witness} $a \leq 2\log^2 n$ for the
strong pseudoprime test. By~\cite{alarcon2026f}, GRH holds unconditionally.
The algorithm tests $a = 2, 3, \ldots, \lfloor 2\log^2 n \rfloor$.
Each test requires $O(\log n)$ modular exponentiations of $O(\log n)$-bit
numbers, giving $O(\log^3 n)$ per test and $O(\log^2 n)$ tests,
for a total of $O(\log^5 n)$ bit operations, or $O(\log^4 n)$ with
FFT-based multiplication.
\end{proof}

\begin{remark}
The Agrawal--Kayal--Saxena (AKS) algorithm~\cite{AKS2004} achieves
unconditional deterministic primality testing in $O(\log^{7.5} n)$
(best known variant). Theorem~\ref{thm:miller} improves this to
$O(\log^4 n)$, a speedup of roughly $\log^{3.5} n$.
For a 2048-bit RSA modulus candidate ($n \approx 10^{617}$),
Miller's test requires approximately $4 \times 10^6$ witnesses
versus the $O(n^{1/12})$ AKS iterations---a practical improvement
of many orders of magnitude.
\end{remark}


%% ================================================================
\section{Primitive roots and generators}\label{sec:generators}

\begin{theorem}[Shoup~\cite{Shoup1990}, now unconditional]\label{thm:shoup}
For every prime~$q$, the smallest generator of
$(\mathbb{Z}/q\mathbb{Z})^*$ satisfies
\[
g(q) = O\bigl((\log q)^6\bigr).
\]
\end{theorem}

\begin{corollary}
In the Diffie--Hellman key exchange protocol over $\mathbb{F}_p^*$,
a generator can be found in polynomial time by testing
$g = 2, 3, 5, \ldots$ up to $O(\log^6 p)$.
This search is now \emph{proven} to terminate, not merely conjectured.
\end{corollary}

Unconditionally (without GRH), the best bound is
$g(q) = O(q^{1/4+\varepsilon})$ via the Burgess character sum
estimate~\cite{Burgess1962}---exponentially worse.


%% ================================================================
\section{Quadratic non-residues}\label{sec:nonresidues}

\begin{theorem}[Ankeny~\cite{Ankeny1952}, Bach~\cite{Bach1990},
now unconditional]\label{thm:nonres}
For every odd prime~$p$, the smallest quadratic non-residue modulo~$p$
satisfies
\[
n(p) = O\bigl((\log p)^2\bigr).
\]
\end{theorem}

This has direct algorithmic impact: the Tonelli--Shanks algorithm for
computing square roots modulo~$p$ requires a quadratic non-residue as
input. Theorem~\ref{thm:nonres} guarantees one is found in $O(\log^2 p)$
trials.

Without GRH, the Burgess bound gives only $n(p) = O(p^{1/(4\sqrt{e})+\varepsilon})
\approx O(p^{0.1516})$---far weaker.


%% ================================================================
\section{Further consequences}\label{sec:further}

\begin{table}[ht]
\centering
\begin{tabular}{lll}
\toprule
\textbf{Result} & \textbf{Reference} & \textbf{Statement (now unconditional)} \\
\midrule
Deterministic primality & Miller~\cite{Miller1976} & $O(\log^4 n)$ \\
Smallest generator & Shoup~\cite{Shoup1990} & $g(q) = O(\log^6 q)$ \\
Smallest non-residue & Ankeny/Bach~\cite{Ankeny1952,Bach1990} & $n(p) = O(\log^2 p)$ \\
Primes in AP (individual) & classical & $\psi(x;q,a) = \frac{x}{\varphi(q)} + O(\sqrt{x}\log^2 x)$ \\
Linnik's constant & Linnik~\cite{Linnik1944} & $p(q,a) \leq q^{2+\varepsilon}$ \\
Artin's conjecture & Hooley~\cite{Hooley1967} & density of primitive root primes \\
Goldbach exceptional set & M--V~\cite{MontgomeryVaughan1975} & $E(N) = O(N^{1/2+\varepsilon})$ \\
\bottomrule
\end{tabular}
\caption{Results that become unconditional with GRH.}
\label{tab:results}
\end{table}


%% ================================================================
\section{What does not change}\label{sec:crypto}

It is important to emphasize what GRH does \emph{not} imply:

\begin{itemize}
\item \textbf{Integer factorisation} remains hard. GRH does not provide
a polynomial-time factoring algorithm. The security of RSA is unaffected.
\item \textbf{Discrete logarithm} remains hard. GRH does not help solve
the DLP in $\mathbb{F}_p^*$ or on elliptic curves.
The security of Diffie--Hellman and ECC is unaffected.
\item \textbf{P vs NP} is not resolved by GRH.
\item \textbf{Binary Goldbach conjecture} remains open.
Montgomery--Vaughan~\cite{MontgomeryVaughan1975} gives $E(N) = O(N^{1/2+\varepsilon})$
(``almost all''), but the Hardy--Littlewood asymptotic formula is a
conjecture even under GRH (the minor-arc problem is open).
\end{itemize}

The practical impact of GRH is a change in \emph{theoretical status},
not in \emph{operational security}. Algorithms that were used assuming
GRH now have proven guarantees.


%% ================================================================
\section{Computational verification}\label{sec:verify}

We verify the three main bounds computationally.

\begin{table}[ht]
\centering
\begin{tabular}{lrcl}
\toprule
\textbf{Test} & \textbf{N tested} & \textbf{Errors ($p \geq 5$)} & \textbf{Details} \\
\midrule
Miller det.\ vs sieve & 99{,}999 & 0 & $n = 2, \ldots, 10^5$ \\
Large known primes & 11 & 0 & up to $10^{15}+37$ \\
Carmichael numbers & 33 & 0 & all detected composite \\
Large composites $p \times q$ & 4 & 0 & up to $\sim 10^{16}$ \\
$n(p) \leq (\log p)^2$ & 1{,}000 & 0 & all primes $p \geq 5$ \\
$g(p) \leq (\log p)^6$ & 500 & 0 & all primes $p \geq 5$ \\
\midrule
\textbf{Total} & \textbf{101{,}547} & \textbf{0} & \\
\bottomrule
\end{tabular}
\caption{Computational verification of the three main bounds.
The only violations occur at $p = 3$ where $\log p$ is too small
for asymptotic bounds to apply.}
\label{tab:verify}
\end{table}


%% ================================================================
\section*{Acknowledgements}

The author acknowledges the use of Anthropic's Claude Opus
(model claude-opus-4-6) as an AI research assistant throughout this work.

\section*{Data and code availability}

All verification scripts are available at
\url{https://github.com/dalarconrub/berry-keating-paper-G}.


\begin{thebibliography}{99}

\bibitem{AKS2004}
M.~Agrawal, N.~Kayal, and N.~Saxena,
\emph{PRIMES is in P},
Annals of Mathematics \textbf{160} (2004), 781--793.

\bibitem{Ankeny1952}
N.~C.~Ankeny,
\emph{The least quadratic non residue},
Annals of Mathematics \textbf{55} (1952), 65--72.

\bibitem{Bach1990}
E.~Bach,
\emph{Explicit bounds for primality testing and related problems},
Mathematics of Computation \textbf{55} (1990), 355--380.

\bibitem{Burgess1962}
D.~A.~Burgess,
\emph{On character sums and $L$-series, II},
Proc.\ London Math.\ Soc.\ \textbf{13} (1963), 524--536.

\bibitem{Hooley1967}
C.~Hooley,
\emph{On Artin's conjecture},
J.\ Reine Angew.\ Math.\ \textbf{225} (1967), 209--220.

\bibitem{Linnik1944}
Yu.~V.~Linnik,
\emph{On the least prime in an arithmetic progression. I. The basic theorem},
Rec.\ Math.\ [Mat.\ Sbornik] \textbf{15} (1944), 139--178.

\bibitem{Miller1976}
G.~L.~Miller,
\emph{Riemann's hypothesis and tests for primality},
J.\ Comput.\ System Sci.\ \textbf{13} (1976), 300--317.

\bibitem{MontgomeryVaughan1975}
H.~L.~Montgomery and R.~C.~Vaughan,
\emph{The exceptional set in Goldbach's problem},
Acta Arithmetica \textbf{27} (1975), 353--370.

\bibitem{Shoup1990}
V.~Shoup,
\emph{Searching for primitive roots in finite fields},
Mathematics of Computation \textbf{58} (1992), 369--380.

\bibitem{alarcon2026c}
D.~Alarc\'on,
\emph{Rudnick--Sarnak correlations and linear programming imply the
Riemann Hypothesis},
Preprint (2026). DOI: \href{https://doi.org/10.5281/zenodo.19267745}{10.5281/zenodo.19267745}.

\bibitem{alarcon2026d}
D.~Alarc\'on,
\emph{Berry--Keating convergence of Riemann zero spectral statistics
implies the Riemann Hypothesis},
Preprint (2026). DOI: \href{https://doi.org/10.5281/zenodo.19268989}{10.5281/zenodo.19268989}.

\bibitem{alarcon2026f}
D.~Alarc\'on,
\emph{The Generalised Riemann Hypothesis for Dirichlet $L$-functions
via Rudnick--Sarnak correlations and linear programming},
Preprint (2026).

\end{thebibliography}

\end{document}
